The error when inferring phylogenies with incipient species by a Birth-Death model

### 2018-04-16 article.tex
Implemented feedback, thanks @rsetienne

From an incipient species tree, we sample a species tree. To do
so, from each species a sub-species must be chosen to represent
the good species as a whole. To clarify, it may be that an
incipient species branched of from a good species. Both of these
subspecies are recognized as the good species, and both can be 
picked as a representative. Here we use three sampling scenario's,
in which we pick the youngest (most recent incipient speciation), 
oldest or random subspecies. Note that only in few 
cases (when speciation completion rate is low) the incipient species
chosen will have an effect on the branch length distribution of species 
tree, as this requires a very specific scenario.

@@

From an alignment, we run a Bayesian analysis and create a posterior, 
using the phylogetic tool BEAST2 \cite{beast2} using the 
pirouette (\cite{pirouette}) package. As our site model, we assume a Jukes-Cantor 
nucleotide substitution model, as we used that in the simulation of the alignment.
For our clock model, we assume a strict clock with the same fixed rate as 
used in the simulation of the alignment. The tree prior assumed is the BD model, 
as this simplification is the goal of this research. 
Additionally, we assume a MRCA prior with a normal distribution
with a mean of the crown age, and a standard deviation of 0.001 time units. 0.001
time units equates to 1000 years, which is the same resolution as which the
alignment is created. The MCMC chain is run to generate 1100 states,
of which the first 10\% (also called the 'burn-in') is removed. Of the remaining
1000 MCMC states, the effective sample size (ESS) of the posterior must at least be 200
for a strong enough inference (\cite{beastbook}). The MCMC sampling interval is 
chosen in such a way that this is the case for all simulations. For simulations
with an ESS below 200, their MCMC sampling interval is doubled up until
the ESS exceeds 200.

We realize there will be an unavoidable bias in our data set: the MCMC
algorithm will produce phylogenies that are representative for the joint
likelihood of (correct) site model, (correct) clock model and 
(incorrect) (Birth-Death) speciation model. This will
warp the distribution of phylogenies somewhat towards the incorrect
speciation model: branch length distributions that are likely in the
Birth-Death model will be relatively overrepresented. We expect
this bias to be systematic, equally affecting distributions
of identical speciation completion rates, and symptomatic
of using contemporary tools.

>>>>

From an incipient species tree, we sample a species tree. To do
so, from each species a sub-species must be chosen to represent
the good species as a whole. To clarify, it may be that an
incipient species branched of from a good species. Both of these
subspecies are recognized as the good species, and both can be 
picked as a representative. Here we use three sampling scenario's,
in which we pick the youngest (most recent incipient speciation), 
oldest or random subspecies. Note that only in few 
cases (when speciation completion rate is low) the incipient species
chosen will have an effect on the branch length distribution of species 
tree. The simplest scenario requires a species, with one good and
one incipient lineage, after which either lineages gives 
rise to a new good species, before the first incipient lineage
completes speciation itself.

@@

From an alignment, we run a Bayesian analysis and create a posterior, 
using the phylogetic tool BEAST2 \cite{beast2} using the 
pirouette (\cite{pirouette}) package. As our site model, we assume a Jukes-Cantor 
nucleotide substitution model, as we used that in the simulation of the alignment.
For our clock model, we assume a strict clock with the same fixed rate as 
used in the simulation of the alignment. The tree prior assumed is the BD model, 
as this simplification is the goal of this research. 
Additionally, we assume a MRCA prior with a normal distribution
with a mean of the crown age, and a standard deviation of 0.001 time units. 0.001
time units equates to 1000 years, which is the same resolution as which the
alignment is created. The MCMC chain is run to generate 1100 states,
of which the first 10\% (also called the 'burn-in') is removed. Of the remaining
1000 MCMC states [NOTE: Why 1000? Why not 250?], 
the effective sample size (ESS) of the posterior must at least be 200
for a strong enough inference (\cite{beastbook}). An ESS can be increased by increasing
the number of samples or decreasing the autocorrelation between samples. 
Would the ESS be less than 200, we decrease autocorrelation by doubling 
the MCMC sampling interval of that simulation, until the ESS exceeds 200.
